---
title: "Notes on rrdb"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Notes on rrdb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
                    collapse = TRUE,
                    comment = "#>"
                  )

```

# Usage

The three function in the packages allow for the creation and manipulation of
the database file.

First load the library and create some test data

```{r setup}
library(rrdb)
library(xts)

ts <- seq( as.POSIXct("1990-01-01",tz="UTC"),as.POSIXct("1990-02-01",tz="UTC"),by=900 )
D <- xts( 1:length(ts), order.by=ts )
```

Note that time step of the data is 900 seconds and that since there is
potentially an observation at 1970-01-01 00:00:00 there is no offset on the
origin of the timestep sequence.

First let us create a database just smaller then the data

```{r create_database}
## use a tempory file
fn <- tempfile()
on.exit( unlink(fn) )

tz_step <- 900
n <- nrow(D)-20
create_db(fn,tz_step,n,ncol(D))
```

Next write the first `n` data points to the database

```{r update_database}
update_db(fn,D[1:n,])
```

To read from the database we supply start and end times for the data. All data
in this period will be returned as an xts object.

```{r read_database}
head( read_db(fn,index(D)[10],index(D)[n-5]) )
```

Updating the database with more data then it can hold will result in the
most recent data being stored.

```{r update_excess}
update_db(fn,D)
head( read_db(fn,index(D)[n],index(D)[nrow(D)]) )
```

# Implementation notes

The functions are written in pure R. They are not especially efficent
implimentation, particularly the calls to `seek` when writing and reading data
are always made from the start of the file. Given both `xts` objects and the
database are ordered this could be greatly improved.

Using `clean=FALSE` when wrting tot he database can be much faster, however
this removes all checks that stop newer data being overwritten by older values
that may be passed in. With the check in place this results in an error.

